---
title: "HUDM6052 Psychometric II Homework_03"
author: "Chenguang Pan (cp3280@tc.columbia.edu)"
date: "2023-10-26"
output:
  pdf_document:
    toc: false
    toc_depth: 4
    number_sections: false
    keep_tex: true
    highlight: tango
---
```{=latex}
\setcounter{tocdepth}{4}
\tableofcontents
```
```{r setup, include=FALSE}
## Global options
knitr::opts_chunk$set(
  cache = TRUE,
  prompt = TRUE,
  comment = '',
  collapse = TRUE,
  warning = FALSE,
  message = FALSE)
```   

## Q1  

*Using the item parameters given in the Table...*  

**My Solution:**  

```{r}
# write the corresponding information function
iif <- function(theta, a, b){
  # get the logit
  Z <- a*(theta-b)
  # get the probability
  out <- 1/(1 + exp(-Z))
  info <- out*(1-out)*(a^2)
  return(info)
}

# set the ability range
theta <- seq(-3,3, by=0.5)
# set the item parameters
a <- c(2, 1.5, 1.5, 1.5, 2)
b <- c(-1, -0.5, 0, 0.5, 1)
# define the color
color_set <- c("red", "green", "blue","violet","black")

# plot the item information function
info_out <- iif(theta, a[1], b[1])
# create a vector to sum up all the information function
test_info <- info_out

# initialize the plot by plotting the first item
plot(theta, info_out,type = "l", col=color_set[1],
     main = "Item/Test Information and SEs",
     xlab = "Ability", ylab = "information",
     ylim = c(0,3))
grid()

# plot the rest item using a for loop
for (i in 2:5) {
  info_out_i <- iif(theta,a[i],b[i])
  lines(theta, info_out_i, type ="l", col=color_set[i])
  test_info <- test_info +info_out_i
}

# draw the test information function
lines(theta, test_info, type = "l", col="gray")
# plot the SE 
SE <- c()
for (j in 1:length(theta)) {
  se_j <- 1/sqrt(sum(iif(theta[j],a,b)))
  SE[j] <- se_j
}
lines(theta, SE, type = "l", col="pink")

# add a legend
legend('topright',inset=0.05,c("item 1","item 2","item 3","item 4","item 5",
                              "test information","SE"),
       lty=1,col=c("red", "green","blue","violet","black","gray","pink"),
       title="Line Type", cex = 0.5)
```  

  
  
## Q2.a  

*a. For each of the six items given in the Table below...*  

**My Solution:**  
The maximum value of a 3PL's information function is at $$\theta_{max}=\beta_i + \frac{1}{a_i}log[\frac{1+\sqrt{1+8c_i}}{2}].$$  
Therefore, I write a function to get the optimal $\theta_{max}$ first. And then send this value together with the parameters into the information function for 3PL model to get the results.
```{r}
# write a function to get the theta_max
get_theta <- function(a,b,c){
  out <- b + log(0.5+0.5*sqrt(1+8*c))/a
  return(out)
}

# write the 3pl information function
iif_3pl <- function(theta, a, b, c){
  z <- a*(theta - b)
  p <- c + (1-c)/(1+exp(-z))
  p_star <- 1/(1 + exp(-z))
  I <- (a^2)*p*(1-p)*(p_star/p)^2
  return(I)
}
```  
Next, plug the given paramters into the functions to get the results.
```{r}
# load the given parameters vectors
b <- c(1,1,1,-1.5,-0.5,0.5)
a <- c(1.8,0.8,1.8,1.8,1.2,0.4)
c <- c(0,0,0.25,0,0.1,0.15)

# using a for loop to get all the required values
theta_vec <- c()
info_vec <- c()
for (i in 1:length(b)) {
  # get the optimal theta value
  theta_max <- get_theta(a = a[i],b = b[i],c = c[i])
  theta_vec[i] <- theta_max
  # get the maximum value of information of item i
  info_max <- iif_3pl(theta = theta_max,a = a[i],b = b[i],c = c[i])
  info_vec[i] <- info_max
}

# Merge all the values as a dataframe
df_out <- data.frame(
  item = seq(1,6),
  theta_max = theta_vec,
  info_max = info_vec
)

df_out
```  
The maximum values of information and corresponding $\theta s$ are shown at end of the code chunk above.  


## Q2.b  

*b. Which item would you choose to make up a two-item...*  

**My Solution:**  
I will choose the `item 1` and `item 2` to make a two-item test since they have the maximum information at the $\theta = 1.0$, which means this test can more accurately measure this given test-taker's ability. The test information at $\theta = 1.0$ is $$0.81+0.16=0.97.$$  

## Q3.a  

*a. Determine the standard error of the estimate...*  

**My Solution:**  
```{r}
# load the given paramters
a <- c(1,1,2,2)
b <- c(0,1,1,1.5)
theta <- 1.5

# using the information function created in the Q1 to get the 
# vector of information at the theta=1.5 for each item
info_vec <- iif(theta=1.5, a, b)

# get the SE
SE_j <- 1/sqrt(sum(info_vec))
SE_j
```  
Therefore, the SE for the test-taker with estimated trait $\theta=1.5$ is `r round(SE_j, 3)`.  
  
## Q3.a  

*b. Construct a 95% confidence interval for...*  

**My Solution:**  
The 95% confidence interval of the estimated $\theta=1.5$ is 
$$95\%CI = 1.5\pm 1.96\times 0.679.$$  
Therefore, $95\%CI$ is $[.169, 2.831].$  



